
We propose model-based derivative-free trust-region algorithms for constrained optimization problems involving black-box functions.
Our algorithms are specifically focused on problems for which infeasible points do not provide function values.
The algorithms work by approximating the objective function and each constraint function by polynomial model functions 
constructed by interpolating over a feasible set of evaluated sample points.
The fact that function values cannot be obtained outside of the feasible region mandates that 
only feasible sample points can be used to construct the model functions.
This restriction poses interesting challenges for ensuring that the sample set is well-poised, 
which means that relative positions of the sample points ensure that the interpolated model functions will be
sufficiently accurate approximations of the true functions.
To address these challenges, we propose two strategies for constructing sample sets that ensure good geometry for the sample set: 
an ellipsoidal trust region strategy and a polyhedral trust region strategy.   

Chapter 3 of the thesis is focused on the special case of linear constraints.
For this case, we present several methods for constructing ellipsoidal sample regions within the intersection of the feasible region and an outer trust region.
By adapting a well-known model-improvement algorithm for selecting sample points,
we show how to construct a well-poised sample set over this ellipsoidal region.
This result is used to show that under reasonable assumptions,  
the iterates produced by our trust region algorithm have criticality measures that converge to zero.
This guarantees, in particular, that any accumulation point of the iterates satisfies the first order optimality conditions.
In addition to the ellipsoidal trust region strategy, we also develop a polyhedral trust region strategy.
For this strategy, we propose a modification to the model improvement algorithm aimed at constructing a well-poised sample set over 
a polyhedral trust region.

The ideas developed for the linear case are built upon in Chapter 4, which addresses the case of nonlinear black-box constraints.
A particular challenge, in this case, is the fact that there is no way to ensure that infeasible function evaluations are never attempted.
However, our goal is to minimize instances when this occurs.
Toward that end, we propose a methodology that constructs second-order cones for each nearly-active constraint 
and then constructs an ellipsoidal sample trust region within the intersection of these cones.
We show that under reasonable assumptions, this sample trust region is guaranteed to be feasible when the trust region is sufficiently small.
Using this result, we prove that the criticality measure for the iterates generated by our trust region algorithm converges to zero.
Numerical results are presented comparing the performance of our algorithm to several other algorithms for constrained derivative-free optimization.
 
