\documentclass[11pt,letterpaper]{thesis2}
\input{thesis_preamble}
\input{thesis_definitions}
\input{thesis_environment}
%=======================================
% Steve's definitions for marking up
%=======================================

\begin{document}
\paragraph{Questions}
\begin{enumerate}
\item For the polyhedral trust region approach, do you use an $L_2$-ball or an $L_{\infty}$-ball for the outer trust region?
\end{enumerate}

\begin{enumerate}
\item Redo .bib keys to something meaningful.   I use the following convention:   <author1 last name>.<author2 last name>.<ea (if more than 2 authors)>:<first meaningful word of title>.  For example:
the paper by   Conejo, Karas, Pedroso, Ribeiro, and Sachine titled``Global Convergence of Trust-region Algorithms for Convex Constrained Minimization Without Derivatives'', would have .bibtex key
Conejo.Karas.ea:global  (instead of Conejo:2013:GCT:2620806.2621814, which is what you currently have).

\item The model improvement algorithm given in the background section as well as the modified algorithms given in chapter 3 need to be restated using the pivot polynomials $u_i$ instead of the basis polynomials $\phi_i$.

\item Sort out the role of $\Delta$ in the model-improvement algorithm.  I believe the error bounds assume that $\Delta = \max \|y^i-y^0\|$.

\item Change $\mu^k$ to $\mu^{(k)}$ throughout?
\item Change $P^k$ to $P^{(k)}$?
\item $\sigma(Q^{(k)})$ versus $\kappa(Q^{(k)})$.

\item Give a clear description (or algorithm) for how to construct a suitable ellipsoid for the linear case.   Do this before proving anything.  Then prove that the resulting ellipsoid is suitable.
\end{enumerate}
\end{document}
